---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mpaviewer

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/dbca-wa/mpaviewer/workflows/R-CMD-check/badge.svg)](https://github.com/dbca-wa/mpaviewer/actions)
[![docker](https://github.com/dbca-wa/mpaviewer/actions/workflows/docker.yaml/badge.svg)](https://github.com/dbca-wa/mpaviewer/actions/workflows/docker.yaml)
[![pkgdown](https://github.com/dbca-wa/mpaviewer/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/dbca-wa/mpaviewer/actions/workflows/pkgdown.yaml)
<!-- badges: end -->

The goal of mpaviewer is to visualise marine monitoring data to DBCA staff.

## Architecture
* Data is loaded from a locally saved `.rds` file, or freshly loaded from
  Google Drive or the DBCA data catalogue (and then saved as `.rds`).
  The locally saved file lives on a persistent volume.

## Data
The dashboard runs off a single `mpa_data.rds` R save file which is produced
from a range of source data files (CSV, TXT, SHP).

### Current data ETL
Archived on the DBCA data catalogue as dataset [mpaviewer](https://data.dbca.wa.gov.au/dataset/mpaviewer) are:
* A link to a Google Drive folder which contains a copy of source data files
  in the folder/file/data structure expected by `mpaviewer::generate_data()`.
* Links to the deployed dashboard and this GitHub code repository
* The mpaviewer data file `mpa_data.rds` at a specific resource ID expected by
  `mpaviewer::download_data()`.

#### Update source data  
Data owners can update the Google Drive folder with new data.

#### Generate dashboard data
Whenever the source data was updated, an analyst must download the contents
of the Google Drive folder "mpaviewer" into the directory `inst/data` of a local 
clone of the mpaviewer codebase and run `mpaviewer::generate_data()`, then upload
the data file `mpa_data.rds` to the DBCA data catalogue.

#### Update the dashboard with new data
A maintainer accesses the running `mpaviewer` Docker container through a shell, 
runs R, and executes `mpaviewer::download_data(data_dir="/app/inst/data")`.

### Ideal data ETL
A second Docker container should run `mpaviewer::download_data(data_dir="/app/inst/data")` as a cronjob, following
[this](https://github.com/dbca-wa/etlTurtleNesting/tree/master/cron) example.

The container could also run other R scripts, e.g. scripts downloading source 
data from the DBCA data catalogue, then running `mpaviewer::generate_data()`
on the downloaded copies. The downloading, processing, and uploading of data
can be split out into another R package or remain in mpaviewer.

Data owners could further automate their data pipelines by scripting the upload 
of their source data to the data catalogue, e.g. using R.

## Development
`mpaviewer` is an RShiny app using 
[Shiny modules](https://shiny.rstudio.com/articles/modules.html), 
developed using [`golem`](https://mastering-shiny.org/scaling-modules.html).

The commands to develop new functionality are in `dev/02_dev.R`.

## Release
Once app and Docker image work, create a new version, tag, and push the tag.
```{r eval=FALSE}
styler::style_pkg()
spelling::spell_check_package()
spelling::update_wordlist()

# Code and docs tested, working, committed
usethis::use_version(which="patch")
usethis::use_version(which="minor")
usethis::use_version(which="major")
usethis::edit_file("NEWS.md")

# Document to load new package version. Git commit, tag, and push.
devtools::document()
v <- packageVersion("mpaviewer")
system(glue::glue("git tag -a v{v} -m 'v{v}'"))
system(glue::glue("git push && git push --tags"))
```

## Deployment
GitHub Actions are configured to build a new Docker image for every tag starting
with "v", such as "v1.0.0".
The images are pushed to the GitHub container registry `ghcr.io`.
The DBCA maintainer of `mpaviewer` then updates the tag number (e.g. `1.0.0`)
in the image run by Kubernetes, which will trigger a hot reload of the image.

The commands to deploy a new version are in `dev/03_deploy.R`.
